---
title: "Code Review Demo: Extracting metadata"
author: "Christopher Brown"
date: "1/25/2021 (revised)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

This is a brief example with simulated (though realistic) sample names using `tidyr::extract` to then generate several metadata columns. Note that the `remove` parameter can be set to `FALSE` to retain the input column, which consists of the sample names in this example.

```{r metadata_extraction}
(sample_names <- paste0(paste0("LMB21_Feces_D", c(1,4,7,10,14)), "_M", rep(c(1:5), each=5)))

(meta_df <- sample_names %>%
  data.frame(Sample_name = .) %>%
  tidyr::extract(col = Sample_name, into = c("Experiment", "Tissue", "Day", "Mouse"), 
          regex = "([[:alnum:]]+)_([[:alnum:]]+)_([[:alnum:]]+)_([[:alnum:]]+)", remove = FALSE))
```

What if the metadata is not uniformly structured? If the patterns found within a sample name don't match the expected regex capture groups, **NA** will be returned for all extaction columns.

```{r uneven_sample_names}
(sample_names_uneven <- c(sample_names, paste0("AE_", 1:2)))

# Note: this operation will result in the AE rows having NA for all fields
(meta_df2 <- sample_names_uneven %>%
  data.frame(Sample_name = .) %>%
  tidyr::extract(col = Sample_name, into = c("Experiment", "Tissue", "Day", "Mouse"), 
          regex = "([[:alnum:]]+)_([[:alnum:]]+)_([[:alnum:]]+)_([[:alnum:]]+)", remove = FALSE))
```

Another approach could be to split out control samples, create necessary variables for each group of samples, and then join the resulting dataframes. In this case a full join is necessary so that all observations and variables are kept in the combined dataframe.

```{r multiple_extractions}
(tissues <- dplyr::filter(meta_df2, Experiment == "LMB21")$Sample_name)
(controls <- dplyr::filter(meta_df2, is.na(Experiment))$Sample_name)
  
(tissue_df <- tissues %>%
  data.frame(Sample_name = .) %>%
  tidyr::extract(col = Sample_name, into = c("Experiment", "Tissue", "Day", "Mouse"), 
          regex = "([[:alnum:]]+)_([[:alnum:]]+)_([[:alnum:]]+)_([[:alnum:]]+)", remove = FALSE))

(control_df <- controls %>%
  data.frame(Sample_name = .) %>%
  tidyr::extract(col = Sample_name, into = c("Replicate"), 
          regex = "_([[:alnum:]]+)", remove = FALSE))

(combined_df <- full_join(tissue_df, control_df, by = "Sample_name"))
```


